New lower bounds for Schur and weak Schur numbers

### 2021-10-13 article/schur.tex
modifs ws + les there is / are arrivent

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They 
extended a sum-free partition vertically by repeating it. Next, they used another sum-free partition to color the other half according 
to the line number, as in \hyperref[figure:1]{Figure 1}. We give an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction and with an S-template which is not included in Abbott and Hanson's construction. In both cases, 
the special color is grey.

@@

\begin{figure}[H]
\begin{center}
\textbf{S-template construction}
\setlength{\tabcolsep}{1.5ex}

@@

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of a template-based construction}
\label{figure:2}

@@

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), we write \(x = (\alpha - 1) q + u\) for certain integers \(\alpha \in \mathbb{N}^*\) and 
	\(u \in [\![1,q]\!]\). Obviously, this decomposition unique. Coefficients \(\alpha\) and \(u\) may be interpreted as the row index of \(x\) and 
	as line and column coordinates of \(x\) in the template. A new coloring is defined as:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]

@@

We produced S-templates using a SAT solver \cite{Lingeling2017}, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing us with the largest value of 
\((a, b)\) (in the lexicographic order). In order to reduce the search space when the number of colors exceeded 
five we 
only looked for symmetric S-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}) and 
(\ref{S(n+5)}) are listed in Appendix \ref{S-templates}.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
Appendix \ref{S-templates}). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.

>>>>

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They 
extended a sum-free partition vertically by repeating it. Next, they used another sum-free partition to color the other half according 
to the line number, as in \hyperref[figure:1]{Figure 1}. We give an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction (\hyperref[figure:1]{Figure 1}) and with an S-template which is not included in Abbott and Hanson's construction 
(\hyperref[figure:2]{Figure 2}). In both cases, the special color is grey.

@@

\begin{figure}[H]
\begin{center}
\textbf{S-template construction}
\label{figure:2}
\setlength{\tabcolsep}{1.5ex}

@@

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of a template-based construction}

@@

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), we write \(x = (\alpha - 1) q + u\) for certain integers \(\alpha \in \mathbb{N}^*\) 
	and \(u \in [\![1,q]\!]\). Obviously, this decomposition unique. Coefficients \(\alpha\) and \(u\) may be interpreted as line 
	and column coordinates of \(x\) in the template. We define a coloring \(h\) of \([\![1,pq+m_{n+1}-1]\!]\) using \( [\![1,n+k]\!]\) colors:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]

@@

We produced S-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bound on 
\(S^+\) and inequalities of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing us with 
the largest value of \((a, b)\) (in the lexicographic order). In order to reduce the search space when the number 
of colors exceeded five we only looked for symmetric S-templates, we assumed that the special color was the last 
color to appear and we constrained the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT 
problem can be found in \cite{Heule2017}.

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}) and 
(\ref{S(n+5)}) are listed in \ref{S-templates}.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
\ref{S-templates}). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.
